2-D Photosynthesis Model

### 2023-11-13 ms/si.qmd
final draft of ms/si.pdf

We model leaf photosynthesis using a two-dimensional porous medium approximation. The model is solved using a finite element method (FEM) using the `steady.2d()` function in the *R* package **rootSolve** version `r packageVersion("rootSolve")` [@soetaert_practical_2009]. Table S1 is a glossary model terms and symbols.

@@

We assume that the leaf is a homogenous 2-D medium. The mesophyll is $T_\text{leaf}$ thick and the stomata are regularly spaced apart by distance $U$ on both ab- and adaxial surfaces. In this scenario, we assume that the stomata on each surface are precisely offset from each other by distance $U/2$. This minimizes the average distance between any point in the mesophyll and its nearest stomate. Because of the regular spacing, we only need to model the region between a stomate on surface and the next stomate on the other surface (Fig. S1). The rest of the mesophyll will be the same because of symmetry.

@@

parms |>
  # filter(!calculated) |>
  mutate(Value = ifelse(
    is.na(default_value),
    "$-$",
    glue::glue("${x}$", x = scientize(default_value))
  ),
  units = ifelse(
    is.na(units) | units == "NA",
    "$-$",
    units
  )) |>
  select(Name = name, Symbol = symbol, Value, Units = units, Notes = notes) |>
  knitr::kable(booktabs = TRUE, escape = FALSE, longtable = TRUE)
  # Option 1
  # gt::gt() |>
  # gt::fmt_markdown()

@@

$$D_\text{e} \nabla^2 C_\text{ias} = D_\text{e} \bigg(\frac{\partial^2 C_\text{ias}}{\partial x^2} + \frac{\partial^2 C_\text{ias}}{\partial z^2}\bigg) = -f_\text{liq}$$ {#eq-2d_pm_flux}

@@

@@

is the effective diffusivity of CO$_2$ though a porous medium composed of an intercellular airspace with a porosity ($\varphi$; m$^3$ airspace m$^{-3}$ leaf) and tortuosity ($\tau$; m m$^{-1}$). $D_\text{c}$ is the diffusion coefficient (m s$^{-1}$) for CO$_2$ in the intercellular airspace, $C_\text{ias}$ is the \[CO$_2$\] (mol m$^{-3}$) at horizontal positions $x$ and depth $z$ in the intercellular airspace, $f_\text{liq}$ is the volumetric rate of CO$_2$ diffusion from the intercellular airspace into the chloroplast stroma (mol m$^{-3}$ s$^{-1}$), $r_\text{c}$ is the volumetric rate of ribulose 1,5-bisphosphate (RuBP) carboxylation (mol m$^{-3}$ s$^{-1}$), $r_\text{d}$ is the volumetric respiration rate (mol m$^{-3}$ s$^{-1}$), and $r_\text{p}$ is the volumetric photorespiration rate by Rubisco (mol m$^{-3}$ s$^{-1}$). Following @earles_excess_2017, $r_\text{d}$ is assumed constant per stroma surface area (Table S1) and $r_\text{p}$ is a function of carboxylation ($r_\text{c}$) and $C_\text{liq}$:

$$ r_\text{p} = r_\text{c} \frac{\Gamma^*}{C_\text{liq}} $$ {#eq-rp}

Carboxylation rate is the minimum of the Rubisco-limited ($w_\text{c}$) or RuBP-regeneration limited ($w_\text{j}$) carboxylation rate:

$$ r_\text{c} = \text{min}(w_\text{c}, w_\text{j}) $$ {#eq-rc}

@@

$k_\text{c}$ is the catalytic rate of Rubisco (m$^{-1}$) and $K_\text{m}$ is effective Michaelis-Menten constant for Rubisco (mol m$^{-3}$). Following @earles_excess_2017, we assumed the relative concentration of Rubisco follows that of @nishio_carbon_1993, but scaled such that the bulk leaf Rubisco concentration integrates to $X_\text{c}$ described in Table S1. We estimated a continuous function describing the relative Rubisco profile as a function of leaf depth using a generalized additive model with the `gam()` function in *R* package **mgcv** version `r packageVersion("mgcv")` [@wood_generalized_2017].

@@

$$ j_\text{e} = \text{min}(j_\text{max}, j_\infty) $$ {#eq-je}

The local $j_\text{max}$ follows the same profile as Rubisco as a function of leaf depth and is scaled so that it integrates to $J_\text{max}$ on a leaf-area basis [@earles_excess_2017]. Potential e$^{-}$ transport is assumed proportional to the local chloroplast concentration so that it integrates to $J_\infty$ on a leaf-area basis [@earles_excess_2017], where:

$$ J_\infty = I_0 \alpha \beta \phi_\text{PSII}.$$ {#eq-Jinf}

@@

WORKING HERE

The local chlorophyll concentration (SYMBOL) is a function of leaf depth following Equation XX from @borsuk_spatial_2019:

The volumetric rate of CO$_2$ diffusion from the intercellular airspace into the chloroplast stroma, $f_\text{liq}$, is:

$$f_\text{liq} = \frac{g_\text{liq} (C_\text{liq} - C_\text{ias})}{T_\text{leaf} / S_\text{m}}$$ {#eq-fliq2}

where $g_\text{liq}$ is the CO$_2$ conductance from the intercellular airspace into the chloroplast stroma (m s$^{-1}$), $C_\text{liq}$ (mol m$^{-3}$) is the \[CO$_2$\] in the stroma, and $S_\text{m}$ is leaf surface area-to-mesophyll surface area ratio. Noting that $g_\text{liq}$ is conductance per m$^2$ of stroma, this means the length scale to divide by should be the inverse of stroma area per unit bulk leaf volume, i.e. $1/[S_\text{c} (1 / T_\text{leaf})] = T_\text{leaf} / S_\text{c}$. For simplicity, we assume that the entire mesophyll surface area is lined with chloroplasts, hence $S_\text{m} = S_\text{c}$.

I calculated assimilation and respiration the same way as @earles_excess_2017 using the standard C\$\_3\$ biochemical model.

The boundary conditions are that the CO$_2$ concentration in the substomatal cavity is constant at $C_\text{stom}$. The fluxes on the left and right sides are 0 because of symmetry.

@@

extra words

@@

>>>>

We modeled leaf photosynthesis using a two-dimensional porous medium approximation. The `steady.2d()` function in the *R* package **rootSolve** version `r packageVersion("rootSolve")` [@soetaert_practical_2009] solves the model using a finite element method (FEM). The 2-D leaf profile is $n_x$ elements long and $n_z$ elements deep with square elements of area $t_\text{elem}^2$. In all cases, we set $t_\text{elem} = 1~\mu \text{m}$. Table S1 is a glossary model terms and symbols.

@@

We assume that the leaf is a homogenous 2-D medium. The mesophyll is $T_\text{leaf}$ thick and the stomata are regularly spaced apart by distance $U$ on both ab- and adaxial surfaces. In this scenario, we assume that the stomata on each surface are precisely offset from each other by distance $U/2$. This minimizes the average distance between any point in the mesophyll and its nearest stomate. Because of the regular spacing, we only need to model the region between a stomate on surface and the next stomate on the other surface (Fig. S1). The rest of the mesophyll will be the same because of symmetry. This allowed us to set the boundary fluxes on the left and right sides of the leaf profile to 0.

@@

@@

$$D_\text{e} \nabla^2 C_\text{ias} = D_\text{e} \bigg(\frac{\partial^2 C_\text{ias}}{\partial x^2} + \frac{\partial^2 C_\text{ias}}{\partial z^2}\bigg) = -f_\text{liq},$$ {#eq-2d_pm_flux}

@@

and

@@

is the effective diffusivity of CO$_2$ though a porous medium composed of an intercellular airspace with a porosity ($\varphi$; m$^3$ airspace m$^{-3}$ leaf) and tortuosity ($\tau$; m m$^{-1}$). We assume the palisade ($\varphi_\text{pal}$) is less porous than the spongy ($\varphi_\text{spg}$) mesophyll (Table S1). $D_\text{c}$ is the diffusion coefficient (m s$^{-1}$) for CO$_2$ in the intercellular airspace, $C_\text{ias}$ is the \[CO$_2$\] (mol m$^{-3}$) at horizontal positions $x$ and depth $z$ in the intercellular airspace, $f_\text{liq}$ is the volumetric rate of CO$_2$ diffusion from the intercellular airspace into the chloroplast stroma (mol m$^{-3}$ s$^{-1}$), $r_\text{c}$ is the volumetric rate of ribulose 1,5-bisphosphate (RuBP) carboxylation (mol m$^{-3}$ s$^{-1}$), $r_\text{d}$ is the volumetric respiration rate (mol m$^{-3}$ s$^{-1}$), and $r_\text{p}$ is the volumetric photorespiration rate by Rubisco (mol m$^{-3}$ s$^{-1}$). Following @earles_excess_2017, $r_\text{d}$ is assumed constant per stroma surface area (Table S1) and $r_\text{p}$ is a function of carboxylation ($r_\text{c}$) and $C_\text{liq}$:

$$ r_\text{p} = r_\text{c} \frac{\Gamma^*}{C_\text{liq}}. $$ {#eq-rp}

Carboxylation rate is the minimum of the Rubisco-limited ($w_\text{c}$) or RuBP-regeneration limited ($w_\text{j}$) carboxylation rates:

$$ r_\text{c} = \text{min}(w_\text{c}, w_\text{j}) \frac{S_\text{m}}{T_\text{leaf}} V_\text{strom}, $$ {#eq-rc}

@@

Multiplying by $\frac{S_\text{m}}{T_\text{leaf}} V_\text{strom}$ converts carboxylation from per area to per stroma volume units. $k_\text{c}$ is the catalytic rate of Rubisco (m$^{-1}$) and $K_\text{m}$ is effective Michaelis-Menten constant for Rubisco (mol m$^{-3}$). Following @earles_excess_2017, we assumed the relative concentration of Rubisco follows that of @nishio_carbon_1993, but scaled such that the bulk leaf Rubisco concentration integrates to $X_\text{c}$ described in Table S1. We estimated a continuous function describing the relative Rubisco profile as a function of leaf depth using a generalized additive model with the `gam()` function in *R* package **mgcv** version `r packageVersion("mgcv")` [@wood_generalized_2017].

@@

$$ j_\text{e} = \text{min}(j_\text{max}, j_\infty). $$ {#eq-je}

The local $j_\text{max}$ follows the same depth profile as Rubisco and is scaled by local $S_\text{m}$ and $V_\text{strom}$ so that it integrates to $J_\text{max}$ on a leaf-area basis [@earles_excess_2017]:

$$J_\text{max} = \int_0^{T_\text{leaf}} j_{\text{max},i} S_{\text{m},i} V_\text{strom} dz.$$ {#eq-jmax}

Potential e$^{-}$ transport follows the same depth profile as chlorophyll concentration and is scaled by local $S_\text{m}$ and $V_\text{strom}$ so that it integrates to $J_\infty$ on a leaf-area basis [@earles_excess_2017], where:

$$ J_\infty = I_0 \alpha \beta \phi_\text{PSII} = \int_0^{T_\text{leaf}} j_{\infty,i} S_{\text{m},i} V_\text{strom} dz.$$ {#eq-Jinf}

@@

The local chlorophyll concentration is a function of depth described by a quadratic equation [@johnson_leaf_2005]:

$$F_\text{chl} = f(z) = b_{0,\text{chl}} + b_{1,\text{chl}} z + b_{2,\text{chl}} z ^ 2.$$ {#eq-Fchl}

We used generic parameters from @borsuk_spatial_2019, rescaled to relative depth on a 0-1 rather than 0-100 interval.

For mass balance, the CO$_2$ assimilated via carboxylation must be supplied by diffusive flux from the intercellular airspace into the chloroplast stroma. The volumetric rate of CO$_2$ diffusion from the intercellular airspace into the chloroplast stroma, $f_\text{liq}$, is:

$$f_\text{liq} = \frac{g_\text{liq} (C_\text{liq} - C_\text{ias})}{T_\text{leaf} / S_\text{m}},$$ {#eq-fliq2}

where $g_\text{liq}$ is the CO$_2$ conductance from the intercellular airspace into the chloroplast stroma (m s$^{-1}$), $C_\text{liq}$ (mol m$^{-3}$) is the \[CO$_2$\] in the stroma, and $S_\text{m}$ is mesophyll surface area-to-leaf surface area ratio. Noting that $g_\text{liq}$ is conductance per m$^2$ of stroma, this means the length scale to divide by should be the inverse of stroma area per unit bulk leaf volume, i.e. $1/[S_\text{c} (1 / T_\text{leaf})] = T_\text{leaf} / S_\text{c}$. For simplicity, we assume that the entire mesophyll surface area is lined with chloroplasts, hence $S_\text{m} = S_\text{c}$. Following @earles_excess_2017, we assumed greater surface area in the palisade ($S_\text{m,pal}$) than spongy ($S_\text{m,spg}$) mesophyll. The fractions of palisade and spongy mesophyll are $f_\text{pal}$ and $f_\text{spg}$, respectively.

Following @earles_excess_2017, we set boundary conditions at the stomata, $C_\text{stom}$, to be $0.85 \times$ atmospheric CO$_2$ (Table S1). The fluxes on the left and right sides are 0 because of symmetry; the other fluxes on the ab- and adaxial surface are assumed 0 (i.e. the epidermis is impermeable to CO$_2$).

@@

@@

parms |>
  # filter(!calculated) |>
  mutate(Value = case_when(
    is.na(default_value) & !calculated ~ "variable",
    is.na(default_value) & calculated ~ "calculated",
    is.numeric(default_value) ~ glue::glue("${x}$", x = scientize(default_value))
  ),
  units = ifelse(
    is.na(units) | units == "NA",
    "$-$",
    units
  )) |>
  select(Name = name, Symbol = symbol, Value, Units = units, Notes = notes) |>
  knitr::kable(booktabs = TRUE, escape = FALSE, longtable = TRUE)
  # Option 1
  # gt::gt() |>
  # gt::fmt_markdown()

### 2023-11-13 ms/si.tex
final draft of ms/si.pdf

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden]}{\end{tcolorbox}}\fi

We model leaf photosynthesis using a two-dimensional porous medium
approximation. The model is solved using a finite element method (FEM)
using the \texttt{steady.2d()} function in the \emph{R} package
\textbf{rootSolve} version 1.8.2.4 (Soetaert and Herman 2009). Table S1
is a glossary model terms and symbols.

@@

We assume that the leaf is a homogenous 2-D medium. The mesophyll is
\(T_\text{leaf}\) thick and the stomata are regularly spaced apart by
distance \(U\) on both ab- and adaxial surfaces. In this scenario, we
assume that the stomata on each surface are precisely offset from each
other by distance \(U/2\). This minimizes the average distance between
any point in the mesophyll and its nearest stomate. Because of the
regular spacing, we only need to model the region between a stomate on
surface and the next stomate on the other surface (Fig. S1). The rest of
the mesophyll will be the same because of symmetry.

@@

\(f_{z,\text{chl}} = b_{0,\text{chl}} + b_{1,\text{chl}} z + b_{2,\text{chl}} z ^ 2\) \\

@@

\(-29\) & \(-\) & Borsuk and Brodersen (2019);
\(f_{z,\text{chl}} = b_{0,\text{chl}} + b_{1,\text{chl}} z + b_{2,\text{chl}} z ^ 2\) \\

@@

{[}CO\(_2\){]} in intercellular airspace & \(C_\text{ias}\) & \(-\) &
mol m\(^{-3}\) & Equation~\ref{eq-2d_pm_flux} and

@@

{[}CO\(_2\){]} in chloroplast stroma & \(C_\text{liq}\) & \(-\) & mol
m\(^{-3}\) & Equation~\ref{eq-2d_pm_flux} \\

@@

\(D_\text{e}\) & \(-\) & m\(^{2}\) s\(^{-1}\) & Equation~\ref{eq-De} \\

@@

fluoresence & \(F_\text{z,\text{chl}}\) & \(-\) & 1 & Borsuk and
Brodersen (2019);
\(f_{z,\text{chl}} = b_{0,\text{chl}} + b_{0,\text{chl}} z + b_{2,\text{chl}} z ^ 2\) \\

@@

PPFD incident on the leaf surface & \(I_0\) & \(1.50 \times 10^{-3}\) &
mol m\(^{-2}\) s\(^{-1}\) & assumed \\

@@

\(J_\infty\) & \(-\) & mol m\(^{-2}\) leaf s\(^{-1}\) &

@@

basis & \(j_\text{e}\) & \(-\) & mol m\(^{-3}\) stroma s\(^{-1}\) &

@@

basis & \(j_\infty\) & \(-\) & mol m\(^{-3}\) stroma s\(^{-1}\) & see
text \\

@@

& \(j_\text{max}\) & \(-\) & mol m\(^{-3}\) stroma s\(^{-1}\) & see
text \\

@@

Number of elements in \(x\) direction & \(n_x\) & \(100\) & \(-\) &

@@

Number of elements in \(z\) direction & \(n_z\) & \(200\) & \(-\) &

@@

\(\varphi_\text{pal}\) & \(0.1\) & m\(^3\) airspace m\(^{-3}\) leaf &

@@

Volumetric rate of RuBP carboxylation & \(r_\text{c}\) & \(-\) & mol
m\(^{-2}\) stroma s\(^{-1}\) & Equation~\ref{eq-rc} \\

@@

\(-\) & mol m\(^{-2}\) stroma s\(^{-1}\) & Earles et al. (2017);

@@

Leaf surface area-to-mesophyll surface area ratio, palisiade &

@@

Leaf surface area-to-mesophyll surface area ratio spongy &

@@

Leaf thickness & \(T_\text{leaf}\) & \(-\) & m &

@@

Interstomatal distance & \(U\) & \(-\) & m &

@@

Rubisco-limited carboxylation rate & \(w_\text{c}\) & \(-\) & mol

@@

RuBP regeneration-limited carboxylation rate & \(w_\text{j}\) & \(-\) &
mol m\(^{-2}\) stroma s\(^{-1}\) & Equation~\ref{eq-wj} \\

@@

\begin{equation}\protect\hypertarget{eq-2d_pm_flux}{}{D_\text{e} \nabla^2 C_\text{ias} = D_\text{e} \bigg(\frac{\partial^2 C_\text{ias}}{\partial x^2} + \frac{\partial^2 C_\text{ias}}{\partial z^2}\bigg) = -f_\text{liq}}\label{eq-2d_pm_flux}\end{equation}

is the effective diffusivity of CO\(_2\) though a porous medium composed
of an intercellular airspace with a porosity (\(\varphi\); m\(^3\)
airspace m\(^{-3}\) leaf) and tortuosity (\(\tau\); m m\(^{-1}\)).
\(D_\text{c}\) is the diffusion coefficient (m s\(^{-1}\)) for CO\(_2\)
in the intercellular airspace, \(C_\text{ias}\) is the {[}CO\(_2\){]}
(mol m\(^{-3}\)) at horizontal positions \(x\) and depth \(z\) in the
intercellular airspace, \(f_\text{liq}\) is the volumetric rate of
CO\(_2\) diffusion from the intercellular airspace into the chloroplast
stroma (mol m\(^{-3}\) s\(^{-1}\)), \(r_\text{c}\) is the volumetric
rate of ribulose 1,5-bisphosphate (RuBP) carboxylation (mol m\(^{-3}\)
s\(^{-1}\)), \(r_\text{d}\) is the volumetric respiration rate (mol
m\(^{-3}\) s\(^{-1}\)), and \(r_\text{p}\) is the volumetric
photorespiration rate by Rubisco (mol m\(^{-3}\) s\(^{-1}\)). Following
Earles et al. (2017), \(r_\text{d}\) is assumed constant per stroma
surface area (Table S1) and \(r_\text{p}\) is a function of
carboxylation (\(r_\text{c}\)) and \(C_\text{liq}\):

\begin{equation}\protect\hypertarget{eq-rp}{}{ r_\text{p} = r_\text{c} \frac{\Gamma^*}{C_\text{liq}} }\label{eq-rp}\end{equation}

Carboxylation rate is the minimum of the Rubisco-limited
(\(w_\text{c}\)) or RuBP-regeneration limited (\(w_\text{j}\))
carboxylation rate:

\begin{equation}\protect\hypertarget{eq-rc}{}{ r_\text{c} = \text{min}(w_\text{c}, w_\text{j}) }\label{eq-rc}\end{equation}

\(k_\text{c}\) is the catalytic rate of Rubisco (m\(^{-1}\)) and
\(K_\text{m}\) is effective Michaelis-Menten constant for Rubisco (mol
m\(^{-3}\)). Following Earles et al. (2017), we assumed the relative
concentration of Rubisco follows that of Nishio, Sun, and Vogelmann
(1993), but scaled such that the bulk leaf Rubisco concentration
integrates to \(X_\text{c}\) described in Table S1. We estimated a
continuous function describing the relative Rubisco profile as a
function of leaf depth using a generalized additive model with the
\texttt{gam()} function in \emph{R} package \textbf{mgcv} version 1.9.0
(Wood 2017).

\begin{equation}\protect\hypertarget{eq-je}{}{ j_\text{e} = \text{min}(j_\text{max}, j_\infty) }\label{eq-je}\end{equation}

The local \(j_\text{max}\) follows the same profile as Rubisco as a
function of leaf depth and is scaled so that it integrates to
\(J_\text{max}\) on a leaf-area basis (Earles et al. 2017). Potential
e\(^{-}\) transport is assumed proportional to the local chloroplast
concentration so that it integrates to \(J_\infty\) on a leaf-area basis
(Earles et al. 2017), where:

\begin{equation}\protect\hypertarget{eq-Jinf}{}{ J_\infty = I_0 \alpha \beta \phi_\text{PSII}.}\label{eq-Jinf}\end{equation}

The local chlorophyll concentration (SYMBOL) is a function of leaf depth
following Equation XX from Borsuk and Brodersen (2019):

The volumetric rate of CO\(_2\) diffusion from the intercellular
airspace into the chloroplast stroma, \(f_\text{liq}\), is:

\begin{equation}\protect\hypertarget{eq-fliq2}{}{f_\text{liq} = \frac{g_\text{liq} (C_\text{liq} - C_\text{ias})}{T_\text{leaf} / S_\text{m}}}\label{eq-fliq2}\end{equation}

where \(g_\text{liq}\) is the CO\(_2\) conductance from the
intercellular airspace into the chloroplast stroma (m s\(^{-1}\)),
\(C_\text{liq}\) (mol m\(^{-3}\)) is the {[}CO\(_2\){]} in the stroma,
and \(S_\text{m}\) is leaf surface area-to-mesophyll surface area ratio.
Noting that \(g_\text{liq}\) is conductance per m\(^2\) of stroma, this
means the length scale to divide by should be the inverse of stroma area
per unit bulk leaf volume,
i.e.~\(1/[S_\text{c} (1 / T_\text{leaf})] = T_\text{leaf} / S_\text{c}\).
For simplicity, we assume that the entire mesophyll surface area is
lined with chloroplasts, hence \(S_\text{m} = S_\text{c}\).

I calculated assimilation and respiration the same way as Earles et al.
(2017) using the standard C\$\_3\$ biochemical model.

The boundary conditions are that the CO\(_2\) concentration in the
substomatal cavity is constant at \(C_\text{stom}\). The fluxes on the
left and right sides are 0 because of symmetry.

extra words

@@

>>>>

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, sharp corners, breakable, frame hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

We modeled leaf photosynthesis using a two-dimensional porous medium
approximation. The \texttt{steady.2d()} function in the \emph{R} package
\textbf{rootSolve} version 1.8.2.4 (Soetaert and Herman 2009) solves the
model using a finite element method (FEM). The 2-D leaf profile is
\(n_x\) elements long and \(n_z\) elements deep with square elements of
area \(t_\text{elem}^2\). In all cases, we set
\(t_\text{elem} = 1~\mu \text{m}\). Table S1 is a glossary model terms
and symbols.

@@

We assume that the leaf is a homogenous 2-D medium. The mesophyll is
\(T_\text{leaf}\) thick and the stomata are regularly spaced apart by
distance \(U\) on both ab- and adaxial surfaces. In this scenario, we
assume that the stomata on each surface are precisely offset from each
other by distance \(U/2\). This minimizes the average distance between
any point in the mesophyll and its nearest stomate. Because of the
regular spacing, we only need to model the region between a stomate on
surface and the next stomate on the other surface (Fig. S1). The rest of
the mesophyll will be the same because of symmetry. This allowed us to
set the boundary fluxes on the left and right sides of the leaf profile
to 0.

\begin{equation}\protect\hypertarget{eq-2d_pm_flux}{}{D_\text{e} \nabla^2 C_\text{ias} = D_\text{e} \bigg(\frac{\partial^2 C_\text{ias}}{\partial x^2} + \frac{\partial^2 C_\text{ias}}{\partial z^2}\bigg) = -f_\text{liq},}\label{eq-2d_pm_flux}\end{equation}

and

is the effective diffusivity of CO\(_2\) though a porous medium composed
of an intercellular airspace with a porosity (\(\varphi\); m\(^3\)
airspace m\(^{-3}\) leaf) and tortuosity (\(\tau\); m m\(^{-1}\)). We
assume the palisade (\(\varphi_\text{pal}\)) is less porous than the
spongy (\(\varphi_\text{spg}\)) mesophyll (Table S1). \(D_\text{c}\) is
the diffusion coefficient (m s\(^{-1}\)) for CO\(_2\) in the
intercellular airspace, \(C_\text{ias}\) is the {[}CO\(_2\){]} (mol
m\(^{-3}\)) at horizontal positions \(x\) and depth \(z\) in the
intercellular airspace, \(f_\text{liq}\) is the volumetric rate of
CO\(_2\) diffusion from the intercellular airspace into the chloroplast
stroma (mol m\(^{-3}\) s\(^{-1}\)), \(r_\text{c}\) is the volumetric
rate of ribulose 1,5-bisphosphate (RuBP) carboxylation (mol m\(^{-3}\)
s\(^{-1}\)), \(r_\text{d}\) is the volumetric respiration rate (mol
m\(^{-3}\) s\(^{-1}\)), and \(r_\text{p}\) is the volumetric
photorespiration rate by Rubisco (mol m\(^{-3}\) s\(^{-1}\)). Following
Earles et al. (2017), \(r_\text{d}\) is assumed constant per stroma
surface area (Table S1) and \(r_\text{p}\) is a function of
carboxylation (\(r_\text{c}\)) and \(C_\text{liq}\):

\begin{equation}\protect\hypertarget{eq-rp}{}{ r_\text{p} = r_\text{c} \frac{\Gamma^*}{C_\text{liq}}. }\label{eq-rp}\end{equation}

Carboxylation rate is the minimum of the Rubisco-limited
(\(w_\text{c}\)) or RuBP-regeneration limited (\(w_\text{j}\))
carboxylation rates:

\begin{equation}\protect\hypertarget{eq-rc}{}{ r_\text{c} = \text{min}(w_\text{c}, w_\text{j}) \frac{S_\text{m}}{T_\text{leaf}} V_\text{strom}, }\label{eq-rc}\end{equation}

Multiplying by \(\frac{S_\text{m}}{T_\text{leaf}} V_\text{strom}\)
converts carboxylation from per area to per stroma volume units.
\(k_\text{c}\) is the catalytic rate of Rubisco (m\(^{-1}\)) and
\(K_\text{m}\) is effective Michaelis-Menten constant for Rubisco (mol
m\(^{-3}\)). Following Earles et al. (2017), we assumed the relative
concentration of Rubisco follows that of Nishio, Sun, and Vogelmann
(1993), but scaled such that the bulk leaf Rubisco concentration
integrates to \(X_\text{c}\) described in Table S1. We estimated a
continuous function describing the relative Rubisco profile as a
function of leaf depth using a generalized additive model with the
\texttt{gam()} function in \emph{R} package \textbf{mgcv} version 1.9.0
(Wood 2017).

\begin{equation}\protect\hypertarget{eq-je}{}{ j_\text{e} = \text{min}(j_\text{max}, j_\infty). }\label{eq-je}\end{equation}

The local \(j_\text{max}\) follows the same depth profile as Rubisco and
is scaled by local \(S_\text{m}\) and \(V_\text{strom}\) so that it
integrates to \(J_\text{max}\) on a leaf-area basis (Earles et al.
2017):

\begin{equation}\protect\hypertarget{eq-jmax}{}{J_\text{max} = \int_0^{T_\text{leaf}} j_{\text{max},i} S_{\text{m},i} V_\text{strom} dz.}\label{eq-jmax}\end{equation}

Potential e\(^{-}\) transport follows the same depth profile as
chlorophyll concentration and is scaled by local \(S_\text{m}\) and
\(V_\text{strom}\) so that it integrates to \(J_\infty\) on a leaf-area
basis (Earles et al. 2017), where:

\begin{equation}\protect\hypertarget{eq-Jinf}{}{ J_\infty = I_0 \alpha \beta \phi_\text{PSII} = \int_0^{T_\text{leaf}} j_{\infty,i} S_{\text{m},i} V_\text{strom} dz.}\label{eq-Jinf}\end{equation}

The local chlorophyll concentration is a function of depth described by
a quadratic equation (Johnson et al. 2005):

\begin{equation}\protect\hypertarget{eq-Fchl}{}{F_\text{chl} = f(z) = b_{0,\text{chl}} + b_{1,\text{chl}} z + b_{2,\text{chl}} z ^ 2.}\label{eq-Fchl}\end{equation}

We used generic parameters from Borsuk and Brodersen (2019), rescaled to
relative depth on a 0-1 rather than 0-100 interval.

For mass balance, the CO\(_2\) assimilated via carboxylation must be
supplied by diffusive flux from the intercellular airspace into the
chloroplast stroma. The volumetric rate of CO\(_2\) diffusion from the
intercellular airspace into the chloroplast stroma, \(f_\text{liq}\),
is:

\begin{equation}\protect\hypertarget{eq-fliq2}{}{f_\text{liq} = \frac{g_\text{liq} (C_\text{liq} - C_\text{ias})}{T_\text{leaf} / S_\text{m}},}\label{eq-fliq2}\end{equation}

where \(g_\text{liq}\) is the CO\(_2\) conductance from the
intercellular airspace into the chloroplast stroma (m s\(^{-1}\)),
\(C_\text{liq}\) (mol m\(^{-3}\)) is the {[}CO\(_2\){]} in the stroma,
and \(S_\text{m}\) is mesophyll surface area-to-leaf surface area ratio.
Noting that \(g_\text{liq}\) is conductance per m\(^2\) of stroma, this
means the length scale to divide by should be the inverse of stroma area
per unit bulk leaf volume,
i.e.~\(1/[S_\text{c} (1 / T_\text{leaf})] = T_\text{leaf} / S_\text{c}\).
For simplicity, we assume that the entire mesophyll surface area is
lined with chloroplasts, hence \(S_\text{m} = S_\text{c}\). Following
Earles et al. (2017), we assumed greater surface area in the palisade
(\(S_\text{m,pal}\)) than spongy (\(S_\text{m,spg}\)) mesophyll. The
fractions of palisade and spongy mesophyll are \(f_\text{pal}\) and
\(f_\text{spg}\), respectively.

Following Earles et al. (2017), we set boundary conditions at the
stomata, \(C_\text{stom}\), to be \(0.85 \times\) atmospheric CO\(_2\)
(Table S1). The fluxes on the left and right sides are 0 because of
symmetry; the other fluxes on the ab- and adaxial surface are assumed 0
(i.e.~the epidermis is impermeable to CO\(_2\)).

@@

Equation~\ref{eq-Fchl} \\

@@

\(-29\) & \(-\) & Borsuk and Brodersen (2019); Equation~\ref{eq-Fchl} \\

@@

{[}CO\(_2\){]} in intercellular airspace & \(C_\text{ias}\) & calculated
& mol m\(^{-3}\) & Equation~\ref{eq-2d_pm_flux} and

@@

{[}CO\(_2\){]} in chloroplast stroma & \(C_\text{liq}\) & calculated &
mol m\(^{-3}\) & Equation~\ref{eq-2d_pm_flux} \\

@@

\(D_\text{e}\) & calculated & m\(^{2}\) s\(^{-1}\) &
Equation~\ref{eq-De} \\

@@

fluoresence & \(F_\text{chl}\) & calculated & 1 & Borsuk and Brodersen
(2019); Equation~\ref{eq-Fchl} \\

@@

PPFD incident on the leaf surface & \(I_0\) & variable & mol m\(^{-2}\)

@@

\(J_\infty\) & calculated & mol m\(^{-2}\) leaf s\(^{-1}\) &

@@

basis & \(j_\text{e}\) & calculated & mol m\(^{-3}\) stroma s\(^{-1}\) &

@@

basis & \(j_\infty\) & calculated & mol m\(^{-3}\) stroma s\(^{-1}\) &

@@

& \(j_\text{max}\) & calculated & mol m\(^{-3}\) stroma s\(^{-1}\) &
Equation~\ref{eq-jmax} \\

@@

Number of elements in \(x\) direction & \(n_x\) & calculated & \(-\) &

@@

Number of elements in \(z\) direction & \(n_z\) & calculated & \(-\) &

@@

\(\varphi_\text{pal}\) & variable & m\(^3\) airspace m\(^{-3}\) leaf &

@@

Volumetric rate of RuBP carboxylation & \(r_\text{c}\) & calculated &
mol m\(^{-2}\) stroma s\(^{-1}\) & Equation~\ref{eq-rc} \\

@@

calculated & mol m\(^{-2}\) stroma s\(^{-1}\) & Earles et al. (2017);

@@

Mesophyll surface area-to-leaf surface area ratio, palisiade &

@@

Mesophyll surface area-to-leaf surface area ratio, spongy &

@@

Leaf thickness & \(T_\text{leaf}\) & variable & m &

@@

Interstomatal distance & \(U\) & variable & m &

@@

Rubisco-limited carboxylation rate & \(w_\text{c}\) & calculated & mol

@@

RuBP regeneration-limited carboxylation rate & \(w_\text{j}\) &
calculated & mol m\(^{-2}\) stroma s\(^{-1}\) & Equation~\ref{eq-wj} \\

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-johnson_leaf_2005}{}}%
Johnson, Daniel M., William K. Smith, Thomas C. Vogelmann, and Craig R.
Brodersen. 2005. {``Leaf Architecture and Direction of Incident Light
Influence Mesophyll Fluorescence Profiles.''} \emph{American Journal of
Botany} 92 (9): 1425--31. \url{https://doi.org/10.3732/ajb.92.9.1425}.
